Manuscript_draft

### 2021-03-06 manuscript_draft.Rmd
growth dominance statistics

# Create species mean contrasts table
speciesFitM.contrasts<-

@@

>>>>

# Create species mean contrasts table

@@

### Growth dominance statistics

```{r calculate growth dominance}

# Filter only for trees that have measurements in 1992 and 2019 
gd.treeList<- # assign variable
  X %>% 
  filter(TreeID!="New") %>% # remove new trees
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(Status=="Live") %>% # filter for live trees
  group_by(TreeID) %>% 
  summarise(N=n()) %>% 
  filter(N>=2) %>% #filter for trees with at least 2 measurements
  pull(TreeID)

# derive periodic increment (period diameter or basal area increment)
gd.BAInc<-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% gd.treeList) %>% 
  mutate(BA=d2ba(DBH)) %>% # convert DBH to basal area
  dplyr::select(Plot,TreeID,Species,Year,BA) %>% 
  pivot_wider(names_from=Year,values_from=BA) %>% 
  mutate(PBAI_2019=`2019`-`2009`,
         PBAI_2009=`2009`-`1997`,
         PBAI_1997=`1997`-`1994`,
         PBAI_1994=`1994`-`1992`) %>% 
  dplyr::select(Plot,TreeID,Species,starts_with("PBAI_"))

gd.Inc<-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% gd.treeList) %>% 
  left_join(filter(volCoef,VolType=="total"),by="Species") %>% 
  mutate(vol=exp(1)^b0*DBH^b1*Height.Pred^b2) %>% 
  dplyr::select(Plot,TreeID,Species,Year,Vol=vol) %>% 
  pivot_wider(names_from=Year,values_from=Vol) %>%
   mutate(PVI_2019=`2019`-`2009`,
         PVI_2009=`2009`-`1997`,
         PVI_1997=`1997`-`1994`,
         PVI_1994=`1994`-`1992`) %>% 
  dplyr::select(Plot,TreeID,Species,starts_with("PVI_")) %>% 
  inner_join(gd.BAInc,by=c("Plot","TreeID","Species")) %>% 
  pivot_longer(
    cols = -c("Plot","Species","TreeID"),
    names_to = c("Var", "Year"),
    names_sep = "_",
    values_to = "Inc")

gd.prelim<-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% gd.treeList) %>% 
  dplyr::select(Plot,TreeID,Species,Year,DBH.Init=DBH) %>% 
  mutate(Year=fct_recode(Year,tmp_1994="1992",
                         tmp_1997="1994",
                         tmp_2009="1997",
                         tmp_2019="2009")) %>% 
  mutate(Year=str_remove(as.character(Year),pattern="tmp_")) %>% 
  mutate(Year=factor(Year)) %>% 
  mutate(BA.Init=d2ba(DBH.Init)) %>% 
  dplyr::select(Plot:Year,BA.Init)

gd.Dat <-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% gd.treeList) %>% 
  dplyr::select(Plot,TreeID,Species,Year,Status,DBH,Height) %>% 
  left_join(gd.Inc,by=c("Plot","TreeID","Species","Year")) %>% 
  left_join(gd.prelim,by=c("Plot","TreeID","Species","Year")) %>% 
  filter(Status=="Live") %>% 
  drop_na(Inc)

# clean up some objects
  rm(gd.prelim,gd.BAInc,gd.Inc)

# Plots of GD

  ## Basal area
  gd.cumu.Inc<-  
    gd.Dat %>% 
    filter(Var=="PBAI") %>% 
    arrange(Plot,Year,BA.Init) %>% 
    group_by(Plot,Year) %>%

    # CUmulative increment
    mutate(Cumu.Inc=cumsum(Inc)) %>% 
    mutate(Cumu.Inc2=max(Cumu.Inc)) %>% 
    mutate(Cumu.Prop.Inc=Cumu.Inc/Cumu.Inc2) %>% 
    dplyr::select(-Cumu.Inc2) %>%

    # Cumulative stand basal area
    mutate(Cumu.BA=cumsum(BA.Init)) %>% 
    mutate(Cumu.BA2=max(Cumu.BA)) %>% 
    mutate(Cumu.Prop.BA=Cumu.BA/Cumu.BA2) %>% 
    dplyr::select(-Cumu.BA2)

    # Calculate Growth Dominance statistic
    gd.stats<-
      gd.cumu.Inc %>% 
      mutate(GDS=(Cumu.Prop.Inc+lag(Cumu.Prop.Inc,default=0))*(Cumu.Prop.BA-lag(Cumu.Prop.BA,default=0))) %>% 
      ungroup() %>% 
      group_by(Plot,Year) %>% 
      summarise(GDS=1-sum(GDS))

    gd.stats %>% 
      ggplot()+
      aes(x=Year,y=GDS)+
      geom_point()+
      facet_wrap(~Plot)

    ggplot()+
    aes(x=Cumu.Prop.BA,y=Cumu.Prop.Inc)+
    geom_point()+
    geom_line()+
    geom_abline(slope=1,intercept=0,linetype = "dashed")+
    theme(aspect.ratio = 1)+
    facet_wrap(~Year,scales="free_y")
